---
title: "Investigating the Relationship between Different Lifestyle Factors and Mental Health "
author: "470354388"
date: "`r format(Sys.time(), '%d %B, %Y %H:%M')`"
bibliography: [refs/bibliography.bibtex, refs/Packages.bib]
link-citations: true
output: 
  html_document: 
    self_contained: true # Creates a single HTML file as output
    code-tools: true # Includes a menu to download the code file
    code_folding: hide # Code folding; allows you to show/hide code chunks
    code_download: true # Includes a menu to download the code file
    toc: true # (Optional) Creates a table of contents!
    toc_float: true # table of contents at the side
    number_sections: true # (Optional) Puts numbers next to heading/subheadings
    
---

```{r setup, include=FALSE, echo = FALSE,  message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)

knitr::write_bib(c(.packages(),
                   "knitr", "rmarkdown"), "refs/Packages.bib")
```

```{r, echo = FALSE,  message = FALSE}

raw_data = readr::read_tsv("./Survey Data.tsv")
old_names = colnames(raw_data)
new_names = c("timestamp","covid_positive","living_arrangements","height","uni_travel_method","uni_travel_listen","spain_budget","feel_overseas","feel_anxious","study_hrs","read_news","study_load","work","lab_zoom","social_media","gender","sleep_time","wake_time","random_number","steak_preference","dominant_hand","normal_advanced","exercise_hrs","employment_hrs","city","weekly_saving","hourly_plan","weeks_behind","assignment_on_time","used_r_before","team_role","data2x02_hrs","social_media_hrs","uni_year","sport","wam","shoe_size","decade_selection")
# overwrite the old names with the new names:
colnames(raw_data) = new_names


```


# Introduction
@Orygen


Is this a random sample of DATA2X02 students?

What are the potential biases? Which variables are most likely to be subjected to this bias?

Which questions needed improvement to generate useful data (e.g. in terms of the way the question was phrased or response validation)?

This report borrowed formatting and code snippets from  @Garth_Tar

# Results

## Is there a significant difference between the sleep times of students with different emplyment requirements? 


```{r,  message = FALSE}
#Question 3 - sleep time, work - difference between 2 groups (full time study only) 

data_1 <- raw_data %>%
  filter(study_load == 'Full time') %>%
  select(c(work, sleep_time, wake_time)) %>%
  drop_na() %>%
  mutate(x1 = replace(x1, x1 == 2, 99))
  

```

## Does physical activity significantly impact the mental health of students? 

```{r,  message = FALSE}
# Question 1 - sport hours vs mental health - resampling of some kind 

```

```{r,  message = FALSE}
# Question 2 - covid19 and wam - compare - paired t test

```

# References
